---
title: "R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
---

#Conclusion
According to the ROC Curve regarding the validation set, we can see that random forest appears to be the best model to predict diabetes, for this reason we applied the random forest model to the test set in order to predict the presence of the diabetes. 
To sum up: our random forest model take into consideration 4 variables and the prediction will be the average results of 300 decision trees:
```{r}
dataset <- read.csv(file.choose())
datasetT <- read.csv(file.choose())
rownames(dataset) <- dataset$X
dataset <- dataset[2:10]
dataset$Class_variable <- factor(dataset$Class_variable, labels=c('No','Yes'))
rownames(datasetT) <- datasetT$X
datasetT <- datasetT[2:10]
library(randomForest)
rf <- randomForest(dataset$Class_variable~., mtry = 4, data=dataset, ntree=300)
datasetT$Outcome <- predict(rf, newdata = datasetT, type='response')
```
```{r}
head(datasetT)
tail(datasetT)
write.csv(datasetT, 'testoutcome.csv')

```{r}
library(ggplot2)
ggplot(datasetT,aes(x=Outcome))+
  geom_bar(aes(y = (..count..)/sum(..count..)))+
  scale_y_continuous(labels=scales::percent)+ 
  ylab('Percentage')+
  xlab('Diabetes')+
  ggtitle('Prediction on test set')

```
Our test set contains about 60% of patients that didn't have diabetes and about 30% of patients that have got diabetes.
```{r}
library(GGally)
ggpairs(datasetT, aes(colour=Outcome), lower = list(continuous = wrap("points", size=0.6)))
```
From the scatterplot we can see that the conditional analysis of the presence of diabetes in test set almost coincide with the conditional analysis of the train set.

